Safe Transferable Regions

### 2016-10-06 paper/type-inference.tex
constraint solver

\paragraph{Constraint Solver.}
Our algorithm for solving a set of constraints $C$ consists of the following steps:
\begin{enumerate}
\item We use an iterative fixed point (saturation) algorithm to compute
$\saturate{C}$, the set of all constraints logically implied by $C$.
\item Let $\groundc{C}$ denote the subset of all ground constraints
(\ie, constraints without any region variable or predicate variable)
in $\saturate{C}$. We check if $\groundc{C}$ is valid in the theory of
partial-order. If $\groundc{C}$ is valid, the solution $\soln$ is determined
as below. Otherwise, the constraint solver fails, as no valid solution exists.
\begin{itemize}
\item For any predicate variable $\varphi$, $\predSubstFn(\varphi)$ is defined
to be $\conj \{ \pi_1 \outlives \pi_2 \;|\; \pi_1, \pi_2 \in \regionConstants, \isvalid{\varphi}{\pi_1 \outlives \pi_2} \in \saturate{C} \}$.
\item For any region variable $\rho$,  we are guaranteed that there exists a unique
 $\pi \in \regionConstants$  such that $\saturate{C}$ contains some constraint
with a consequent $\rho = \pi$. We define $\regionSubstFn(\rho)$ to be $\pi$.
\end{itemize}
\end{enumerate}

Solving the constraints is an iterative process, consisting of the following
steps, and this process fails at any point if it determines that no solution
exists for the set of constraints:

@@

\item
    A constraint of the form $\isvalid{\varphi}{\rho_i = \rho_j}$ unifies
$\rho_i$ and $\rho_j$ if at least one of them is a region variable.
If $\rho_i$ and $\rho_j$ are distinct region constants, then the constraint
solver fails.

\item
When a region variable $\rho$ is unified with a region constant $\pi$,
if $\pi \in \regionDeltaMap(\rho)$, we update $\regionSubstFn(\rho)$ to be $\pi$,
and replace every occurrence of $\rho$ by $\pi$ in the set of constraints.
If $\pi \not\in \regionDeltaMap(\rho)$, the constraint solver fails.

\item
\label{item:context}
Consider any constraint of the form $\isvalid{\varphi \conj \phictxt}{\pi_i \outlives \pi_j}$,
where $\pi_i$ and $\pi_j$ are both region constants.
If $\{ \pi_i, \pi_j \} \subseteq \predDeltaMap(\varphi)$, then we simply add
$\pi_i \outlives \pi_j$ as an additional conjunct to $\predSubstFn(\varphi)$.
Otherwise, we check if $\isvalid{\phictxt}{\pi_i \outlives \pi_j}$.
If this entailment does not hold, the constraint solver fails, since no valid solution is possible.

\item
  When any outlives constraint $\phi$ is added to $\predSubstFn(\varphi)$,
  then for every existing validity constraint $\isvalid{\varphi' \conj \phictxt}{F(\varphi)}$,
  we produce a new validity constraint $\isvalid{\varphi' \conj \phictxt}{F(\phi)}$. 
\end{enumerate}

>>>>

\paragraph{Constraint Solver.}
Our algorithm for solving a set of constraints $C$ consists of the following steps:
\begin{enumerate}
\item We use an iterative fixed point (saturation) algorithm to compute
$\saturate{C}$, the set of all constraints logically implied by $C$.
\item Let $\groundc{C}$ denote the subset of all ground constraints
(\ie, constraints without any region variable or predicate variable)
in $\saturate{C}$. We check if $\groundc{C}$ is valid in the theory of
partial-order. If $\groundc{C}$ is valid, the solution $\soln$ is determined
as below. Otherwise, the constraint solver fails, as no valid solution exists.
\begin{itemize}
\item For any predicate variable $\varphi$, $\predSubstFn(\varphi)$ is defined
to be $\conj \{ \pi_1 \outlives \pi_2 \;|\; \pi_1, \pi_2 \in \regionConstants, \isvalid{\varphi}{\pi_1 \outlives \pi_2} \in \saturate{C} \}$.
\item For any region variable $\rho$,  the set
 $\{ \pi \in \regionConstants \:|\; \isvalid{\ell}{\rho = \pi} \in \saturate{C} \}$ 
is guarantted to contain a single element $\pi$.
We define $\regionSubstFn(\rho)$ to be $\pi$.
%\item For any region variable $\rho$,  we are guaranteed that there exists a unique
% $\pi \in \regionConstants$  such that $\saturate{C}$ contains some constraint
%with a consequent $\rho = \pi$. We define $\regionSubstFn(\rho)$ to be $\pi$.
\end{itemize}
\end{enumerate}

We compute $\saturate{C}$ by iteratively applying the following rules until a fixed
point is reached:

@@

\item (Initialization) $\isvalid{\ell}{r} \in C \Rightarrow \isvalid{\ell}{r} \in \saturate{C}$

\item (Transitivity)
$\isvalid{\ell}{\rho_1 \outlives \rho_2} \in \saturate{C}$,
$\isvalid{\ell}{\rho_2 \outlives \rho_3} \in \saturate{C}$,
$\Rightarrow$
$\isvalid{\ell}{\rho_1 \outlives \rho_3} \in \saturate{C}$

\item (Substitution)
$\isvalid{\ell}{F(\varphi)} \in \saturate{C}$,
$\isvalid{\varphi}{\phi} \in \saturate{C}$
$\Rightarrow$ $\isvalid{\ell}{F(\phi)} \in \saturate{C}$

\item (Abduction Decomposition)
\label{item:context}
\begin{enumerate}
\item
$\isvalid{\varphi \conj \phictxt}{\pi_i \outlives \pi_j} \in \saturate{C}$,
$\{ \pi_i, \pi_j \} \subseteq \predDeltaMap(\varphi)$
$\Rightarrow$
$\isvalid{\varphi}{\pi_i \outlives \pi_j} \in \saturate{C} $

\item 
$\isvalid{\varphi \conj \phictxt}{\pi_i \outlives \pi_j} \in \saturate{C}$,
$\{ \pi_i, \pi_j \} \subseteq \regionConstants \setminus \predDeltaMap(\varphi)$
$\Rightarrow$
$\isvalid{\phictxt}{\pi_i \outlives \pi_j} \in \saturate{C} $
\end{enumerate}

%\item
%\label{item:context}
%Consider any constraint of the form $\isvalid{\varphi \conj \phictxt}{\pi_i \outlives \pi_j}$,
%where $\pi_i$ and $\pi_j$ are both region constants.
%If $\{ \pi_i, \pi_j \} \subseteq \predDeltaMap(\varphi)$, then we simply add
%$\pi_i \outlives \pi_j$ as an additional conjunct to $\predSubstFn(\varphi)$.
%Otherwise, we check if $\isvalid{\phictxt}{\pi_i \outlives \pi_j}$.
%If this entailment does not hold, the constraint solver fails, since no valid solution is possible.
%
%\item
%    A constraint of the form $\isvalid{\varphi}{\rho_i = \rho_j}$ unifies
%$\rho_i$ and $\rho_j$ if at least one of them is a region variable.
%If $\rho_i$ and $\rho_j$ are distinct region constants, then the constraint
%solver fails.
%
%\item
%When a region variable $\rho$ is unified with a region constant $\pi$,
%if $\pi \in \regionDeltaMap(\rho)$, we update $\regionSubstFn(\rho)$ to be $\pi$,
%and replace every occurrence of $\rho$ by $\pi$ in the set of constraints.
%If $\pi \not\in \regionDeltaMap(\rho)$, the constraint solver fails.
%\item
%  When any outlives constraint $\phi$ is added to $\predSubstFn(\varphi)$,
%  then for every existing validity constraint $\isvalid{\varphi' \conj \phictxt}{F(\varphi)}$,
%  we produce a new validity constraint $\isvalid{\varphi' \conj \phictxt}{F(\phi)}$. 
\end{enumerate}
